---
title: "Autumn Team Tournament"
author: "by NN Sheep"
output: 
  flexdashboard::flex_dashboard:
    theme: 
      version: 4
      bootswatch: lux
    vertical_layout: fill
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)
library(tidyverse)
library(knitr)
library(flexdashboard)

# load in data
raw <- readxl::read_excel('autumn_team.xlsx') 
```

```{r}
#### INITIALISATIONS ####

# initialise matchups 
# (rows 1:2=matchup1, 3:4=matchup2, 5:6=matchup3, cols=day)
h2h <- tibble(
  c(6,6,6,6,6),
  c(2,3,5,1,4),
  c(5,5,3,5,5),
  c(3,4,1,2,1),
  c(4,1,2,3,3),
  c(1,2,4,4,2)
  )

# plot template
theme_blank <- function(){
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_blank())}

# colour palette for conistency
pal <- c('#800000','#767676','#FFA319','#8A9045','#155F83','#C16622')

# function to create df for each day (might be a better solution)
tmp_df <- function(x,d){
  return(filter(x,Day==d) %>% 
           mutate(Team=fct_reorder(Team,-Rank)) %>%
           arrange(Rank))
}

h2h_df <- function(x,d){
  pivot_wider(x, id_cols=h2h, names_from=Result, values_from=Team) %>% 
    select(-h2h) 
}

indiv_df <- function(x,d){
  filter(x,Day==d) %>%
  select(Rank,Player,Time,Team) 
}
```

``` {r}
#### DATA PREP ####

# gather columns of same type of info and add column for day
results <- pivot_longer(raw,-c(1:2), values_transform = as.character) %>%
  # collate cols into one day col and one time col
  separate(name,c('Data','Day'),'_') %>% 
  pivot_wider(names_from = Data, values_from = value) %>%
  mutate(Time=round(as.numeric(Time),2), Day=as.numeric(Day)) %>%
  # sort by lowest time for each day and add a rank column
  arrange(Day,Time) %>%
  mutate(Rank=rep(1:n_distinct(Player),n_distinct(Day)),.before=Player) 
results$Team[results$Team==0] <- NA # change teamless players to NA

# summarise team performance
team_results <- results %>%
  filter(!is.na(Team)) %>% # remove teamless players
  group_by(Day,Team) %>%
  summarise(Total=sum(Time),
            Time=round(mean(Time),2)) %>% # average within teams
  # sort by lowest team time for each day and add rank
  arrange(Day,Time) %>%
  ungroup() %>%
  mutate(Rank=rep(1:n_distinct(Team),n_distinct(Day)),
         dblTeam=Team,.before=Team) 
# recode team numbers for table printing convenience later
team_results$Team <- sub('^', 'Team ', team_results$Team)

# add matchups per day (there is probs a better way to do this)
for(i in 1:5){
  #for each day, extract the matchups from h2h table
  team_results$h2h[team_results$Day==i] <- ifelse(
    team_results$dblTeam[team_results$Day==i] %in% as.numeric(h2h[i,1:2]),1,ifelse(
      team_results$dblTeam[team_results$Day==i] %in% as.numeric(h2h[i,3:4]),2,3))
  
  #add time as a proportion of the leaders for the day 
  team_results$prop[team_results$Day==i] <- team_results$Time[team_results$Day==i] / 
    min(team_results$Time[team_results$Day==i])
  #add time as a zscore
  team_results$z[team_results$Day==i] <- (
    team_results$Time[team_results$Day==i] - mean(team_results$Time[team_results$Day==i])
    ) / sd(
      team_results$Time[team_results$Day==i]
      )

  # add win/loss per matchup per day
  for(j in 1:3){
  team_results$Result[team_results$Day==i&team_results$h2h==j] <- ifelse(
    team_results$Rank[team_results$Day==i&team_results$h2h==j] ==
      min(team_results$Rank[team_results$Day==i&team_results$h2h==j]),'Win','Loss')
  }
}
```

# Day 1
##### **Car: `r raw$Car_1[1]`**
##### **Track: `r raw$Track_1[1]`**
</br>

Column 
--------------
### Head2Head Results {data-height=300}
```{r}
d=1 # set day number
tmp <- tmp_df(team_results,d) # create temporary df for the day
kable(h2h_df(tmp,d)) # overall team rankings
```
### Average Time {data-height=600}
```{r}
# plot team performance
ggplot(tmp, aes(Time,Team,fill=Team)) +
  geom_col(position = 'dodge', show.legend=FALSE) +
  scale_fill_manual(values = pal[rev(tmp$dblTeam)]) +
  coord_cartesian(xlim=c(min(tmp$Time)*.999,max(tmp$Time)**1.002)) +
  theme_blank() + 
  annotate('text', x=rev(tmp$Time)+mean(tmp$Time)/650, y=1:6, 
           label=paste0(as.character(rev(tmp$Time)),'s'))
```

Column 
--------------
### Individual Leaderboard 
```{r}
# Table of overall player rankings
kable(indiv_df(results,d))
```


# Day 2
##### Car: `r raw$Car_2[1]`
##### Track: `r raw$Track_2[1]`
</br>

Column 
--------------
### Head2Head Results {data-height=300}
```{r}
d=2 # set day number
tmp <- tmp_df(team_results,d) # create temporary df for the day
kable(h2h_df(tmp,d)) # overall team rankings
```

### Average Time {data-height=600}
```{r}
# plot team performance
ggplot(tmp, aes(Time,Team,fill=Team)) +
  geom_col(position = 'dodge', show.legend=FALSE) +
  scale_fill_manual(values = pal[rev(tmp$dblTeam)]) +
  coord_cartesian(xlim=c(min(tmp$Time)*.999,max(tmp$Time)**1.002)) +
  theme_blank() + 
  annotate('text', x=rev(tmp$Time)+.025, y=1:6, 
           label=paste0(as.character(rev(tmp$Time)),'s'))
```


Column 
--------------

### Individual Leaderboard 
```{r}
# Table of overall player rankings
kable(indiv_df(results,d))
```


# Day 3

##### Car: `r raw$Car_3[1]`
##### Track: `r raw$Track_3[1]`
</br>

Column 
--------------
### Head2Head Results {data-height=300}
```{r}
d=3 # set day number
tmp <- tmp_df(team_results,d) # create temporary df for the day
kable(h2h_df(tmp,d)) # overall team rankings
```

### Average Time {data-height=600}
```{r}
# plot team performance
ggplot(tmp, aes(Time,Team,fill=Team)) +
  geom_col(position = 'dodge', show.legend=FALSE) +
  scale_fill_manual(values = pal[rev(tmp$dblTeam)]) +
  coord_cartesian(xlim=c(min(tmp$Time)*.999,max(tmp$Time)**1.002)) +
  theme_blank() + 
  annotate('text', x=rev(tmp$Time)+.045, y=1:6, 
           label=paste0(as.character(rev(tmp$Time)),'s'))
```

Column 
--------------

### Individual Leaderboard 
```{r}
# Table of overall player rankings
kable(indiv_df(results,d))
```


# Day 4 
##### Car: `r raw$Car_4[1]`
##### Track: `r raw$Track_4[1]`
</br>

Column 
--------------
### Head2Head Results {data-height=300}
```{r}
d=4 # set day number
tmp <- tmp_df(team_results,d) # create temporary df for the day
kable(h2h_df(tmp,d)) # overall team rankings
```
### Average Time {data-height=600}
```{r}
# plot team performance
ggplot(tmp, aes(Time,Team,fill=Team)) +
  geom_col(position = 'dodge', show.legend=FALSE) +
  scale_fill_manual(values = pal[rev(tmp$dblTeam)]) +
  coord_cartesian(xlim=c(min(tmp$Time)*.999,max(tmp$Time)*1.002)) +
  theme_blank() + 
  annotate('text', x=rev(tmp$Time)+mean(tmp$Time)/650, y=1:6, 
           label=paste0(as.character(rev(tmp$Time)),'s'))
```

Column 
--------------
### Individual Leaderboard 
```{r}
# Table of overall player rankings
kable(indiv_df(results,d))
```


# Day 5

##### Car: `r raw$Car_5[1]`
##### Track: `r raw$Track_5[1]`
</br>

Column 
--------------
### Head2Head Results {data-height=300}
```{r}
d=5 # set day number
tmp <- tmp_df(team_results,d) # create temporary df for the day
kable(h2h_df(tmp,d)) # overall team rankings
```

### Average Time {data-height=600}
```{r}
# plot team performance
ggplot(tmp, aes(Time,Team,fill=Team)) +
  geom_col(position = 'dodge', show.legend=FALSE) +
  scale_fill_manual(values = pal[rev(tmp$dblTeam)]) +
  coord_cartesian(xlim=c(min(tmp$Time)*.999,max(tmp$Time)*1.002)) +
  theme_blank() + 
  annotate('text', x=rev(tmp$Time)+.05, y=1:6, 
           label=paste0(as.character(rev(tmp$Time)),'s'))
```


Column 
--------------
### Individual Leaderboard 
```{r}
# manually add in tie
kable(indiv_df(results,d))
```


# Final Standings 

Column 
--------------
### Team Standings {data-height=400}
```{r}
kable(
  team_results %>% 
    group_by(Team) %>%
    summarise(`Total Points` = sum(Result=='Win'),
              `Total Time` = sum(Time)) %>%
    arrange(desc(`Total Points`),`Total Time`)
)
```

### Team Performance Overview {data-height=500}
```{r}
# plot team performance
ggplot(team_results, aes(as.factor(Day),prop,fill=Team)) +
  geom_col(position = 'dodge') +
  scale_fill_manual(values = pal) +
  coord_cartesian(ylim=c(min(team_results$prop)*.99,max(team_results$prop)*1.01)) +
  geom_hline(yintercept = 1, linewidth = 1, linetype = 'dashed') +
  theme_minimal() +
  labs(x = 'Day', y="% Slower than Fastest Team") +
  scale_y_continuous(breaks = c(.99,1,1.01,1.02,1.03,1.04),
                     labels = c('','0','1','2','3','4'))
```

Column 
--------------
### Final Leaderboard: Total Time
```{r}
# Table of overall player rankings
kable(
  results %>%
    select(Player,Time,Team) %>% 
    group_by(Player) %>%
    summarise(`Total Time` = sum(Time)) %>%
    arrange(`Total Time`) %>%
    mutate(Rank = 1:nrow(.),.before = Player)
  )
```

Column 
--------------
### Final Leaderboard: Average Placement
```{r}
# Table of overall player rankings
kable(
  results %>%
    select(Player,Rank,Team) %>% 
    group_by(Player) %>%
    summarise(`Average Placement` = mean(Rank)) %>%
    arrange(`Average Placement`) %>%
    mutate(Rank = 1:nrow(.),.before = Player)
  )
```

```{r}
# plot team performance
ggplot(team_results, aes(as.factor(Day),z,fill=Team)) +
  geom_col(position = 'dodge') +
  scale_fill_manual(values = pal) +
  coord_cartesian(ylim=c(min(team_results$z)*.99,max(team_results$z)*1.01)) +
  geom_hline(yintercept = 0, linewidth = 1, linetype = 'dashed') +
  theme_minimal() +
  labs(x = 'Day', y="Time (z)") 
```
